ArKanjo: a tool for detecting function-level Code Duplication in the Linux Kernel

### 2025-06-11 paper/paper.tex
add references

The Linux kernel is a foundational Free/Libre/Open Source Software (FLOSS) project, 
critical to a significant portion of the world's digital infrastructure. Maintaining 
the kernel is an enormous undertaking, involving more than 28 million lines of code 
and contributions from over twenty thousand developers. Within this context, code 
duplication persists as a significant challenge, a known harmful practice that 
negatively affects code readability and increases the likelihood of introducing bugs 
(Hordijk et al., 2009; Hotta et al., 2012). This issue is particularly acute in the 
kernel's device drivers, which comprise over 66\% of the source code (Schmitt, 2022). 
For instance, maintainers of the AMD Display driver have specifically highlighted code 
duplication as a significant impediment to their work.

The detection of code duplication, or code clones, has been a subject of research for 
decades (Jankowitz, 1988). The literature provides a standard taxonomy, classifying 
clones into four types based on their degree of similarity, from identical copies 
(Type-1) to semantically equivalent but textually different fragments (Type-4) (Chen et al., 2022). 
Various detection methodologies have been developed, including textual, token-based, 
tree-based, and graph-based approaches (Chen et al., 2022). These have culminated in 
state-of-the-art techniques, such as the graph-based work by Liu et al. (2023). 
However, the primary focus of such academic work remains on determining if a given 
pair of code artifacts are duplicates, rather than providing a scalable method to scan 
an entire codebase for actionable results. Conversely, existing free software tools explored 
in practice often lack this sophistication, such as relying on primitive textual matching 
that is insufficient for the complexity and scale of the kernel.

@@

For the text similarity method, we treat the source code files as text and apply the
TF-IDF vector embedding method implemented by the Gensim library (ADD REF),
then compute cosine similarity as the similarity metric. This method was chosen for 
its claimed performance, programming language independence, and the fact that it does 
not require compilable code, which is expected in the
temporary codebase, as it does not contain complete code artifacts.
(SHOULD I ADD THE FOLLOWING? THERE IS LITTLE SPACE:) 
Plates implemented this code duplication detection method and released it under
the MIT license(ADD REF). We adapted his implementation to fit our tool’s expected
input and output formats.

For the difference method, 
we implemented a method that considered the number of exactly equal lines between two functions.
For two functions \textit{function1} and \textit{function2}, we compute the similarity as the 
ratio of duplicated lines between the two functions by the total number of lines in both functions.
This code duplication method is considerably simpler and more explainable than the text similarity method, 
as the metric is the ratio of common lines between the two functions. To compute the number of equal lines, we use
the \textit{diff} command built-in in the Linux environment (ADD REF).

@@

To validate the tool against the BigCloneBench dataset (Svajlenko et al., 2014), we followed
the methodology presentedy by by (Liu et al., 2023), we sampled 20,000 pairs from each clone
type, adding 20,000 non-duplicate pairs as negative samples. We applied the same sampling
approach to our tool to ensure a fair evaluation.
Unlike state-of-art methods, our tool does not distinguish between clone types and
identifies duplications at the function level rather than the file level, which is typical in
state-of-the-art tools. Therefore, we adapted the metric calculation method for evaluation
purposes. Specifically, we considered every pair of functions with a similarity metric
equal to or greater than a threshold 𝑋 as duplicates and marked the corresponding file
pairs. A correctly identified duplication pair counts as accurate for its clone type, while
an incorrectly inferred pair is considered incorrect across all clone types.
To understand the impact of varying the similarity threshold, we evaluated our tool using
different threshold values: 30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%. We then analyzed
and discussed the results.

@@

The first reason may be the limitations and known issues with BigCloneBench, as
highlighted by Krinke and Ragkhitwetsagul (Krinke and Ragkhitwetsagul, 2022). The
second reason could relate to the nature of the AMD Display driver, where each code artifact
shares semantic meaning. In contrast, BigCloneBench comprises self-contained code
artifacts without shared semantics. Given that our tool relies on a text-based code clone
detection approach, it may naturally perform worse on the AMD Display driver than on the
BigCloneBench dataset.

@@

The student efforts showed that newcomers could use the tool to make effective contributions. 
Of the 12 student patches, three were accepted, three were refused, and six required follow-up 
fixes based on maintainer feedback. To remove the duplicated code, seven student groups used the 
Parameterize Method (ADD REF) and two used the Extract Method (ADD REF), which are simple refactoring methods. 
This result corroborates that ArKanjo finds actionable duplications and that newcomers can 
successfully contribute patches to the kernel.

@@

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bibliografia}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

 \bibitem{IEEEhowto:kopka}
 H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
   0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

>>>>

The Linux kernel is a foundational Free/Libre/Open Source Software (FLOSS) project, 
critical to a significant portion of the world's digital infrastructure. Maintaining 
the kernel is an enormous undertaking, involving more than 28 million lines of code 
and contributions from over twenty thousand developers. Within this context, code 
duplication persists as a significant challenge, a known harmful practice that 
negatively affects code readability and increases the likelihood of introducing bugs 
\cite{harmone,harmtwo}. This issue is particularly acute in the 
kernel's device drivers, which comprise over 66\% of the source code \cite{marcelo}. 
For instance, maintainers of the AMD Display driver have specifically highlighted code 
duplication as a significant impediment to their work.

The detection of code duplication, or code clones, has been a subject of research for 
decades (Jankowitz, 1988). The literature provides a standard taxonomy, classifying 
clones into four types based on their degree of similarity, from identical copies 
(Type-1) to semantically equivalent but textually different fragments (Type-4) \cite{litreview}. 
Various detection methodologies have been developed, including textual, token-based, 
tree-based, and graph-based approaches \cite{litreview}. These have culminated in 
state-of-the-art techniques, such as the graph-based work by Liu et al \cite{tailor}. 
However, the primary focus of such academic work remains on determining if a given 
pair of code artifacts are duplicates, rather than providing a scalable method to scan 
an entire codebase for actionable results. Conversely, existing free software tools explored 
in practice often lack this sophistication, such as relying on primitive textual matching 
that is insufficient for the complexity and scale of the kernel.

@@

For the text similarity method, we treat the source code files as text and apply the
TF-IDF vector embedding method implemented by the Gensim library \cite{gensim},
then compute cosine similarity as the similarity metric. This method was chosen for 
its claimed performance, programming language independence, and the fact that it does 
not require compilable code, which is expected in the
temporary codebase, as it does not contain complete code artifacts.

For the difference method, 
we implemented a method that considered the number of exactly equal lines between two functions.
For two functions \textit{function1} and \textit{function2}, we compute the similarity as the 
ratio of duplicated lines between the two functions by the total number of lines in both functions.
This code duplication method is considerably simpler and more explainable than the text similarity method, 
as the metric is the ratio of common lines between the two functions. To compute the number of equal lines, we use
the \textit{diff} command built-in in the Linux environment.

@@

To validate the tool against the BigCloneBench dataset \cite{bigclonebench}, we followed
the methodology presentedy by Liu et al. \cite{tailor}, we sampled 20,000 pairs from each clone
type, adding 20,000 non-duplicate pairs as negative samples. We applied the same sampling
approach to our tool to ensure a fair evaluation.
Unlike state-of-art methods, our tool does not distinguish between clone types and
identifies duplications at the function level rather than the file level, which is typical in
state-of-the-art tools. Therefore, we adapted the metric calculation method for evaluation
purposes. Specifically, we considered every pair of functions with a similarity metric
equal to or greater than a threshold X as duplicates and marked the corresponding file
pairs. A correctly identified duplication pair counts as accurate for its clone type, while
an incorrectly inferred pair is considered incorrect across all clone types.
To understand the impact of varying the similarity threshold, we evaluated our tool using
different threshold values: 30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%. We then analyzed
and discussed the results.

@@

The first reason may be the limitations and known issues with BigCloneBench, as
highlighted by Krinke and Ragkhitwetsagul \cite{bigfail}. The
second reason could relate to the nature of the AMD Display driver, where each code artifact
shares semantic meaning. In contrast, BigCloneBench comprises self-contained code
artifacts without shared semantics. Given that our tool relies on a text-based code clone
detection approach, it may naturally perform worse on the AMD Display driver than on the
BigCloneBench dataset.

@@

The student efforts showed that newcomers could use the tool to make effective contributions. 
Of the 12 student patches, three were accepted, three were refused, and six required follow-up 
fixes based on maintainer feedback. To remove the duplicated code, seven student groups used the 
Parameterize Method and two used the Extract Method \cite{refactorbook}, which are simple refactoring methods. 
This result corroborates that ArKanjo finds actionable duplications and that newcomers can 
successfully contribute patches to the kernel.

@@

% use section* for acknowledgment

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv,referencias}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
% \begin{thebibliography}{1}
%  
%  \bibitem{IEEEhowto:kopka}
%  H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%    0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%  
% \end{thebibliography}
